---
title: "Homework: Simulation"
author: "Firstname Lastname"
output: html_document
---

## Instructions

**Spend no more than 4 hours on writing pseudocode for this assignment**

Read the assignment from start to finish - the simulation design will affect your functions.

Work through this assignment in detailed pseudo-code first. See the hearts.Rmd file for an example. - Put your pseudo-code (or list of steps) in \`\`\` blocks with no {} (this is a generic markup chunk).

Then, *if you have time*, implement the game in R and simulate your question.

You will have an opportunity to submit a link to this assignment with working code for an extra (bonus) homework assignment one week after this assignment is due.

When you are finished with the assignment:

1.  Save the file as 10_Lastname_Firstname.Rmd and compile it
2.  Commit the Rmd file and corresponding html file to your homework git repo
3.  Push the commit to github
4.  Locate the Rmd file on github and click on the SHA link for the file
5.  Paste the url of the resulting page into Canvas to submit your assignment.

Your assignment must compile on a different computer.

## Is it in the cards?

In this task, you will create a simulation of a card game played with a standard deck of 52 cards. You might consider games such as [blackjack](https://bicyclecards.com/how-to-play/blackjack/#filter), [war](https://bicyclecards.com/how-to-play/war/#filter), or [go fish](https://bicyclecards.com/how-to-play/go-fish/#filter). You may choose any game that can be assessed probabilistically -- you may want to avoid games which require a large amount of strategy, skill, and bluffing, such as poker. You may also want to avoid games which depend on pair strategies, such as spades.

This homework file is set up to demonstrate how this would be done with the game of [hearts](https://bicyclecards.com/how-to-play/hearts/#filter). As a result, you may not pick hearts as your game of choice.

You don't need to necessarily simulate the full game mechanism, but you do need to implement enough of the game mechanism that you can identify an interesting probability question in your game and answer the question using a simulation.

### Set up your objects

First, you must set up a data frame describing all possible values of your playing items. For instance, if you are using a standard deck of 52 playing cards, you'll need a data frame which has columns for suit and value, and you may want to make value a factor variable (your call on that one). If your game requires that players hold cards in their hand, you may also want to have a column indicating location (deck, discard, player_x).

    Pseudocode steps

```{r}
# R code goes here - you can add more R chunks if you have multiple steps or want to explain each function after it's written.

```

### Write one or more functions to simulate the basic game mechanisms

For instance, if you are simulating a game where all cards are initially held by players, you need to have a function that randomly assigns 13 cards to each player. You may also want a function that controls the game play and functions for each separate step in game play (some of these functions may need to be called using a loop).

    Pseudocode steps

#### Break your problem down into steps and write functions for each step, putting each step under its own level 4 header (\#\#\#\# at the front)

```{r}
# R code goes here - you can add more R chunks if you have multiple steps or want to explain each function after it's written.

```

#### Full Simulation

Your last function should be something that when called returns information about a fully played game.

```{r}

```

### Identify a probability question you can answer with your simulation

For instance, if you are simulating blackjack, you may want to know what the [expected return on a specific strategy is](https://towardsdatascience.com/python-blackjack-simulator-61b591ffb971) (hint: if you're beating the dealer, your simulation is probably wrong).

### Design a simulation to answer the question you identified

Run your simulation with an increasing number of iterations and create some sort of plot showing your results.

Describe any issues you encountered in your simulation, and how you addressed them (or where the whole thing went off the rails, if you couldn't address everything in time).
